---
title: "README"
author: "Bob Rudis"
date: "August 25, 2014"
output:
  md_document:
    variant: markdown_github
---

`iptools` is a set of tools for a working with IPv4 addresses. The aim is to provide functionality not presently available with any existing R package and to do so with as much speed as possible. To that end, many of the operations are written in `Rcpp` and require installation of the `Boost` libraries. A current, lofty goal is to mimic most of the functionality of the Python `iptools` module and make IP addresses first class R objects.

The package also uses the v1 [GeoLite](http://dev.maxmind.com/geoip/legacy/geolite/) MaxMind library to perform basic geolocation of a given IPv4 address. You must manually install both the maxmind library (`brew install geoip` on OS X, `sudo apt-get install libgeoip-dev` on Ubuntu) and the `GeoLiteCity.dat` http://geolite.maxmind.com/download/geoip/database/GeoLiteCity.dat.gz & `GeoLiteASNum.dat`  http://geolite.maxmind.com/download/geoip/database/GeoLiteASNum.dat.gz files for the geolocation/ASN functions to work. If there's interest in porting to the newer \code{libmaxminddb} library/GeoLite2 format, I'll consider updating the package.

The following functions are implemented:

*Revolver-ish*

- `gethostbyaddr` - Returns all `PTR` records associated with an IPv4 address
- `gethostsbyaddr` - Vectorized version of `gethostbyaddr`
- `gethostbyname` - Returns all `A` records associated with a hostname
- `gethostsbyname` - Vectorized version of `gethostbyname`

*IP int/string conversion*

- `ip2long` - Character (dotted-decimal) IPv4 Address Conversion to long integer
- `long2ip` - Intger IPv4 Address Conversion to Character

*CIDR*

- `cidr_ips` - Converts an IPv4 CIDR (e.g. "192.168.1.0/24") to a vector of individual IPv4 addresses
- `ip_in_cidr` - Test if IPv4 addresses are in a CIDR block
- `cidr_range` - IPv4 CIDR to long integer range

*Validation*

- `validateIP` - Validate IPv4 addresses in dotted-decimal notation
- `validateCIDR` - Validate IPv4 CIDRs in dotted-decimal slash notation

*Geo/ASN Lookup*

- `geoip` - Perform (local) maxmind geolocation on IPv4 addresses (see `?geoip` for details)
- `asnip` - Perform (local) maxmind AS # & org lookup on IPv4 addresses (see `?asnip` for details)

*Testing*

- `randomIPs` - generate a vector of valid, random IPv4 addresses (very helpful for testing)

The following data sets are included:

- `ianaports` - IANA Service Name and Transport Protocol Port Number Registry
- `ianaipv4spar` - IANA IPv4 Special-Purpose Address Registry
- `ianaipv4assignments` - IANA IPv4 Address Space Registry
- `ianarootzonetlds` - IANA Root Zone Database
- `ianaprotocolnumbers` - IANA Protocol Numbers


### Installation

```{r eval=FALSE}
devtools::install_git("https://gitlab.dds.ec/bob.rudis/iptools.git")
```

>NOTE: Under Ubuntu (it probably applies to other variants), this only works with the current version (1.55) of the boost library, which I installed via the [launchpad boost-latest](https://launchpad.net/~boost-latest/+archive/ubuntu/ppa/+packages) package:

    sudo add-apt-repository ppa:boost-latest/ppa
    # sudo apt-get install python-software-properties if "add-apt-repository" is not found
    sudo apt-get update
    sudo apt-get install boost1.55 # might need to use 1.54 on some systems
    
>`homebrew` (OS X) users can do: `brew install boost` and it should `#justwork`.

The first person(s) to get this working under Windows/mingw + boost/Rcpp gets a free copy of [our book](http://dds.ec/amzn)

### Usage

```{r}
library(iptools)

# current verison
packageVersion("iptools")

# lookup google
gethostbyname("google.com")

# lookup apple (in reverse)
gethostbyaddr("17.178.96.59")

# decimal and back
ip2long("17.178.96.59")
long2ip(ip2long("17.178.96.59"))

# checking it twice
validateIP(gethostbyname("google.com"))

validateCIDR("8.0.0.0/8")

# geo
maxmindinit()
set.seed(1492)
geoip(randomIPs(25))

set.seed(100000)
asnip(randomIPs(15))

#cidr
table(ip_in_cidr(cidr_ips("192.168.0.0/23"), "192.168.1.0/24"))
```

### Test Results

```{r}
library(iptools)
library(testthat)

date()

test_dir("tests/")

```
