---
title: "README"
author: "Bob Rudis"
date: "August 7, 2014"
output:
  md_document:
    variant: markdown_github
---

`iptools` is a set of tools for a working with IPv4 addresses. The aim is to provide functionality not presently available with any existing R package and to do so with as much speed as possible. To that end, many of the operations are written in `Rcpp` and require installation of the `Boost` libraries. A current, lofty goal is to mimic most of the functionality of the Python `iptools` module and make IP addresses first class R objects.

The following functions are implemented:

- `gethostbyaddr` - Returns all 'PTR' records associated with an IPv4 address
- `gethostbyname` - Returns all 'A' records associated with a hostname
- `ip2long` - Character (dotted-decimal) IPv4 Address Conversion to long integer
- `iptools` - A package to help perform various tasks with/on IPv4 addresses
- `long2ip` - Intger IPv4 Address Conversion to Character
- `validateIP` - Validate IPv4 addresses in dotted-decimal notation
- `validateCIDR` - Validate IPv4 CIDRs in dotted-decimal slash notation

The following data sets are included:

- `ianaports` - IANA Service Name and Transport Protocol Port Number Registry
- `ianaipv4spar` - IANA IPv4 Special-Purpose Address Registry
- `ianaipv4assignments` - IANA IPv4 Address Space Registry


### Installation

```{r eval=FALSE}
devtools::install_git("https://gitlab.dds.ec/bob.rudis/iptools.git")
```

### Usage

```{r}
library(iptools)

# current verison
packageVersion("iptools")

# lookup google
gethostbyname("google.com")

# lookup apple (in reverse)
gethostbyaddr("17.178.96.59")

# decimal and back
ip2long("17.178.96.59")
long2ip(ip2long("17.178.96.59"))

# checking it twice
validateIP(gethostbyname("google.com"))

validateCIDR("8.0.0.0/8")
```

### Test Results

```{r}
library(iptools)
library(testthat)
test_dir("tests/", lis)

```
